---
title: "Craft Beer Analysis"
author: "Nicole Norelli & Allen Miller"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, error=FALSE)
```

# Craft Beer and Brewery Analysis
## An examination of ABV, IBU, Beer types (particularly IPA & Ale), and geographical trends

## Load necessary libraries and data
```{r}
library(tidyverse)
library(gt)
library(DT)
library(viridis)
library(class)
library(caret)
library(e1071)
breweries = read.csv(file.choose(),header = TRUE) #choose Breweries.csv
beers = read.csv(file.choose(),header = TRUE) #choose Beers.csv
```

## Determine missing values
```{r}
sapply(breweries, function(x) sum(is.na(x)))
sapply(beers, function(x) sum(is.na(x)))
```
# Missing Values:
### There are 1005 (41.7%) missing IBU values and 62 (2.6%) missing ABV values
### We will determine if imputing missing values improves any machine learning techniques we apply.

# Number of Breweries per State
## Create table and heatmap
```{r}
stateCount = count(breweries,State)
stateCount$State = trimws(stateCount$State, which="both") 
lookup = data.frame(abb = state.abb, State = state.name) 
stateCountAbb = stateCount
colnames(stateCountAbb)[1] = "abb"
stateCountFullNames = full_join(stateCountAbb,lookup, by="abb") 
stateCountFullNames$State[8] = "Washington, D.C."
stateCount %>% arrange(desc(n)) %>% 
  gt() %>% tab_header(title="Number of Breweries in Each State")
datatable(stateCount, rownames=F, filter="top",caption="Number of Breweries per State",options=list(pageLength=10,scrollX=T))
states <- map_data("state")
stateCountMap = stateCountFullNames
stateCountMap$region <- tolower(stateCountMap$State)
map.df <- merge(states,stateCountMap, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=n))+
  geom_path()+ 
  scale_fill_viridis(option="D", name="Breweries") + ggtitle("Breweries by State")+
  coord_map()
```

# Breweries per State:
## Colorado has the most with 47, followed by California with 39.
## All states and Washington, D.C. have breweries.

# Merge breweries and beer datasets
```{r}
colnames(beers)[5] = "Brew_ID"
breweries2 = full_join(breweries, beers, by = "Brew_ID")
colnames(breweries2)[2] = "Brewery_Name"
colnames(breweries2)[5] = "Beer_Name"
breweries2$State = trimws(breweries2$State, which="both")
head(breweries2,6)
tail(breweries2,6)
```

# Compute median ABV and IBU for each state
## Create bar charts
```{r}
by_state = group_by(breweries2,State)
medianIBU = summarise(by_state, IBU=median(IBU,na.rm=TRUE))
medianABV = summarise(by_state, ABV=median(ABV,na.rm=TRUE))
medianIBU <- filter(medianIBU, !is.na(IBU))
medianABV <- filter(medianABV, !is.na(ABV))
medianIBU %>% arrange(desc(IBU)) %>% gt() %>% tab_header(title="Median IBU by State")
medianABV %>% arrange(desc(ABV)) %>% gt() %>% tab_header(title="Median ABV by State")
medianIBU %>% ggplot(aes(x= fct_reorder(State,IBU), y=IBU, fill=IBU)) + geom_bar(stat="identity") +
  geom_text(aes(State,IBU+1, label=IBU,fill=NULL),data=medianIBU,size=2) + 
  scale_fill_viridis(option = "D") +
  ggtitle("Median IBU by State") + xlab("State") + coord_flip() +
  theme(axis.text=element_text(size=6))
medianABV %>% ggplot(aes(x= fct_reorder(State,ABV), y=ABV, fill=ABV)) + geom_bar(stat="identity") +
  geom_text(aes(State,ABV+.0015, label=ABV,fill=NULL),data=medianABV,size=2) + 
  scale_fill_viridis(option = "D") + ggtitle("Median ABV by State") + xlab("State") + coord_flip() +
  theme(axis.text=element_text(size=6))
```

# Median ABV by State:
## Ranges from 0.04 (UT) to 0.0625 (KY)
# Median IBU by State:
## Ranges from 19 (Wisconsin) to 61 (Maine)

# Identify state with maximum ABV beer and state with maximum IBU beer
```{r}
invisible(which.max(breweries2$ABV))
breweries2[375,]
invisible(which.max(breweries2$IBU))
breweries2[1857,]
```
# State with maximum ABV beer:
## Colorado (Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale) ABV: 0.128
# State with maximum IBU beer:
## Oregon (Bitter Bitch Imperial IPA) IBU: 138

# Summary statistics and distribution of ABV
```{r}
summary(breweries2$ABV)
sd(breweries2$ABV, na.rm=T)
breweries2 %>% ggplot(aes(x=ABV)) + geom_histogram( color = "white", fill = "#7F57BF") + 
  geom_density(alpha = .6, fill = "white") +
  ggtitle("Distribution of ABV")
```

# Summary Statistics:
## Min: 0.001   Median: 0.056   Max: 0.128    Standard Deviation: 0.0135
# Distribution:
## Right-skewed

# Relationship between bitterness and alcohol content: scatterplot to visualize
```{r}
breweries2 %>% ggplot(aes(x=ABV, y=IBU, color=IBU, alpha=ABV)) + geom_point(position="jitter") +
  geom_smooth() + ggtitle("ABV and IBU") + 
  scale_color_viridis(option="D", name="IBU", direction = -1)
```

# Higher IBU appears associated with higher ABV
## Possible explanation:  High IBU beers have more hops and are more bitter.
## To balance bitterness, more malt(suger) is added.
## This sugar and yeast produce higher ABV.

# Investigate difference with respect to IBU and ABV between IPAs and other Ales
## Replace missing IBU values with median IBU of each style
## Make dataframe with only Ales and IPAs and standardize values
## Find optimal number for nearest neighbor classification
```{r}
breweriesFact = breweries2
breweriesFact$Style = factor(breweriesFact$Style)
by_style = group_by(breweriesFact,Style)
medianIBUbyStyle = summarise(by_style, IBU=median(IBU,na.rm=TRUE))
medianIBUbyStyleFiltered = medianIBUbyStyle %>% filter(!is.na(IBU)) %>% filter(Style != "")
breweriesNoIBU = breweries2 %>% filter(is.na(IBU))
FillNAs = merge(breweriesNoIBU, medianIBUbyStyleFiltered, by="Style")
FillNAs$IBU = FillNAs$IBU.y
FillNAs = subset(FillNAs, select = -c(IBU.y, IBU.x))
breweriesPartial = breweries2 %>% filter(!is.na(IBU))
IBUImpute = rbind(breweriesPartial, FillNAs)
AleOrIPAFull = IBUImpute %>% 
   filter(str_detect(IBUImpute$Style, "\\bAle\\b|\\bIPA\\b"))
AleOrIPAFull = AleOrIPAFull %>% mutate(AleType = str_detect(Style, "\\bIPA\\b"))
AleOrIPAFull$AleType[AleOrIPAFull$AleType==TRUE] = "IPA"
AleOrIPAFull$AleType[AleOrIPAFull$AleType==FALSE] = "Ale" 
AleOrIPAFull$AleType = factor(AleOrIPAFull$AleType)
ZAleOrIPAFull = data.frame(ZABV = scale(AleOrIPAFull$ABV), ZIBU = scale(AleOrIPAFull$IBU), AleType = AleOrIPAFull$AleType)
ZAleOrIPAFull = ZAleOrIPAFull %>% filter(!is.na(ZABV))
ZAleOrIPAFull %>% ggplot(aes(x = ZABV, y=ZIBU ,color = AleType)) + geom_point() +
  xlab("Standardized ABV") + ylab("Standardized IBU") + 
  ggtitle("Standardized ABV and IBU for Ale and IPA (with imputed IBU values)") +
  scale_color_viridis_d(begin=.25, end = .75)
classifications = knn.cv(ZAleOrIPAFull[,c(1,2)], ZAleOrIPAFull$AleType,prob=TRUE, k=29)
confusionMatrix(table(classifications,ZAleOrIPAFull$AleType)) 
```
# Ale is associated with lower ABV and IBU compared to IPA
## Using the 29 nearest data points, our model uses ABV and IBU to predict IPA vs. Ale with 89% accuracy.

# Regional trends regarding ABV and IBU
## Using 4 US Census regions, examine ABV and IBU
```{r}
State <- c("AK", "WA", "OR", "CA", "ID", "NV", "UT", "AZ", "MT", "WY", "CO", "NM", "HI",
           "ND", "SD", "NE,", "KS", "MN", "IA", "MO", "WI", "IL", "IN", "MI", "OH", 
           "TX", "OK", "AR", "LA", "MS", "TN", "KY", "WV", "AL", "FL", "GA", "SC", "NC", "VA", "DC", "MD", "DE", 
           "PA", "NY", "VT", "NH", "ME", "MA", "CT", "RI", "NJ")
regions <- c("West", "West", "West", "West", "West", "West", "West", "West", "West", "West", "West", "West", "West", 
             "Mid West", "Mid West", "Mid West", "Mid West", "Mid West", "Mid West", "Mid West", "Mid West", "Mid West", 
             "Mid West", "Mid West", "Mid West", 
             "South", "South", "South", "South", "South", "South", "South", "South", "South", "South", "South", "South", 
             "South", "South", "South", "South", "South", 
             "North East", "North East", "North East", "North East", "North East", "North East", "North East", "North East", 
             "North East")
statesRegions <- data.frame(State, regions)
breweries3 <- merge(breweries2, statesRegions, by = "State")
by_region = group_by(breweries3,regions)
medianRegionIBU = summarise(by_region, IBU=median(IBU,na.rm=TRUE))
medianRegionABV = summarise(by_region, ABV=median(ABV,na.rm=TRUE))
medianRegionIBU %>% ggplot(aes(x= fct_reorder(regions,IBU), y=IBU, fill=IBU)) + geom_bar(stat="identity") +
  geom_text(aes(regions,IBU+1, label=IBU,fill=NULL),data=medianRegionIBU) + scale_fill_viridis(option = "D") +
  ggtitle("Median IBU by Census Region") + xlab("Region")
medianRegionABV %>% ggplot(aes(x= fct_reorder(regions,ABV), y=ABV, fill=ABV)) + geom_bar(stat="identity") +
  geom_text(aes(regions,ABV+.0015, label=ABV,fill=NULL),data=medianRegionABV) + scale_fill_viridis(option = "D") +
  ggtitle("Median ABV by Census Region") + xlab("Region")
```

# The West Region had the highest median IBU level. 
## Given that Oregon is part of the West, and had the beer with the single highest IBU level, it is unsuprising that the West is the highest.

# The West and Mid West both have the highest median ABV levels
## There is very little discrepentcy between all four regions and more data could change this distribution

# Regional trends and major beer categories
## Group beer styles into seven major groups: IPA/Pale Ale, Ale, Pilsner, Lager, Stout, Wheat, and Other
## Examine category by census region
```{r}
Styles = breweriesFact %>% mutate(Style, Category = ifelse(grepl("Stout", Style), "Stout",
                                                     ifelse(grepl("Pilsner", Style), "Pilsner",
                                                      ifelse(grepl("Pilsener", Style), "Pilsner",
                                                      ifelse(grepl("IPA",Style), "IPA or Pale Ale",
                                                      ifelse(grepl("Pale Ale",Style), "IPA or Pale Ale",
                                                      ifelse(grepl("Tripel",Style), "IPA or Pale Ale",
                                                      ifelse(grepl("Lager",Style),  "Lager",
                                                      ifelse(grepl("Marzen",Style), "Lager",
                                                      ifelse(grepl("Bock",Style), "Lager",
                                                      ifelse(grepl("Doppelbock",Style), "Lager",
                                                      ifelse(grepl("Wheat",Style), "Wheat",
                                                      ifelse(grepl("Hefeweizen", Style), "Wheat",
                                                      ifelse(grepl("Witbier", Style), "Wheat",
                                                      ifelse(grepl("Weissbier", Style), "Wheat",
                                                      ifelse(grepl("Dunkelweizen",Style), "Wheat",
                                                      ifelse(grepl("Kölsch", Style), "Ale",
                                                      ifelse(grepl("Altbier",Style), "Ale",
                                                      ifelse(grepl("Dubbel",Style), "Ale",
                                                      ifelse(grepl("Ale",Style), "Ale","Other"))))))))))))))))))))
Styles %>% count(Category) %>% arrange(desc(n)) %>% gt() %>% tab_header(title="Totals per Category")
breweries4 = merge(Styles, statesRegions, by = "State")
breweries4$Category = factor(breweries4$Category)
cat_by_region = group_by(breweries4, regions)
CatCountRegion = cat_by_region %>% count(Category)
CatCountRegion %>% ggplot(aes(x= fct_reorder(regions,n), y=n, fill=fct_relevel(Category, c("IPA or Pale Ale","Stout","Pilsner","Ale","Other","Lager","Wheat")))) +
  geom_bar(stat="identity", position = "fill") + scale_fill_viridis(option = "D", discrete=T, name="Category\nDescending IBU", direction=-1) +
  ggtitle("Categories by Census Region") + xlab("Region") + ylab("Proportion")
```

# IPA and Ale are the most common of the seven categories in all regions
## Proportion of each type of craft beer produced by region is similar

# Examine major beer groups by state
```{r}
state_grp = group_by(Styles,State)
CatCountbyState = state_grp %>% count(Category)
CatCountbyState = CatCountbyState %>% arrange (State, desc(n))
datatable(CatCountbyState, rownames=F, filter="top",caption="Category Count by State",options=list(pageLength=10,scrollX=T))
TopCatbyState = CatCountbyState %>% group_by(State) %>% filter(row_number()==1)
TopCatbyState$Category = factor(TopCatbyState$Category)
lookup = data.frame(abb = state.abb, State = state.name) 
colnames(TopCatbyState)[1] = "abb"
TopCatFullNames = full_join(TopCatbyState,lookup, by="abb") 
TopCatFullNames$State[8] = "district of columbia"
states <- map_data("state")
CatCountMap = TopCatFullNames
CatCountMap$region <- tolower(CatCountMap$State)
map.df <- merge(states,CatCountMap, by="region", all.x=T)
map.df <- map.df[order(map.df$order),] 
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Category))+
  geom_path()+ 
  scale_fill_viridis(discrete = T,option="D") + ggtitle("Most Common Craft Beers by State")+
  coord_map()
```

# IPA is the most common craft beer type throughout in 37 states and District of Columbia.
## 13 states had Ale as the most common type of craft beer produced.
## Only one state (WI) had a type of beer other than Ale or IPA as the most common beer type in production.
## There seems to be a general demand for IPA and Pale Ale, particularly on the east and west coasts
